---
title: "Toxic Chemical Release: A Pollution Outlook"
author: "Joanna Pittari & Cedar Callahan"
output: html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(forcats)
```

## 1. Introduction
This data comes from the Environmental Protection Agency's (EPA) Toxics Release Inventory (TRI) which is collected every year. The raw data from the EPA included 119 variables detailing facility information and location, toxin types, regulation and oversight, methods of disposal, and amount of toxins released. We will be analyzing data from 2011 to 2021 for trends in pollution, frequency of chemical use by chemical category (metal, carcinogen, pbt, pfas, etc), method of chemical disposal, location (county, state, and if a facility is on tribal land), industry sector, and which chemicals are most common in each industry sector. Additionally, we will examine whether chemicals are included in the Clean Air Act (regulates how much can be released) and if there is any correlation between industry sector and amount of Clean Air Act chemicals used.  

## 2. Data

Before we could begin any sort of analysis, we had to join our datasets and narrow down our variables. Additionally, in the original data each column began with a number, which we found challenging to work with in preliminary graphs. This problem was solved by renaming the variables in a way that we felt was logical and easier to work with. Below is the process of creating, combining, narrowing, and simplifying the data.  

```{r data-set-assignment-2011, warning = FALSE}
library(readr)
my_variables <- c("1. YEAR", "4. FACILITY NAME", "5. STREET ADDRESS", "6. CITY", "7. COUNTY", "8. ST", "9. ZIP", "10. BIA", "11. TRIBE", "12. LATITUDE", "13. LONGITUDE", "15. PARENT CO NAME", "17. STANDARD PARENT CO NAME", "18. FEDERAL FACILITY", "20. INDUSTRY SECTOR", "34. CHEMICAL", "35. ELEMENTAL METAL INCLUDED", "39. CLEAN AIR ACT CHEMICAL", "40. CLASSIFICATION", "41. METAL", "42. METAL CATEGORY", "43. CARCINOGEN", "44. PBT", "45. PFAS", "46. FORM TYPE", "47. UNIT OF MEASURE", "48. 5.1 - FUGITIVE AIR", "49. 5.2 - STACK AIR", "50. 5.3 - WATER", "51. 5.4 - UNDERGROUND", "52. 5.4.1 - UNDERGROUND CL I", "53. 5.4.2 - UNDERGROUND C II-V", "54. 5.5.1 - LANDFILLS", "57. 5.5.2 - LAND TREATMENT", "58. 5.5.3 - SURFACE IMPNDMNT", "61. 5.5.4 - OTHER DISPOSAL", "62. ON-SITE RELEASE TOTAL", "63. 6.1 - POTW - TRNS RLSE", "64. 6.1 - POTW - TRNS TRT", "65. POTW - TOTAL TRANSFERS", "85. OFF-SITE RELEASE TOTAL", "91. OFF-SITE RECYCLED TOTAL", "94. OFF-SITE ENERGY RECOVERY T", "101. OFF-SITE TREATED TOTAL", "103. 6.2 - TOTAL TRANSFER", "104. TOTAL RELEASES", "105. 8.1 - RELEASES", "106. 8.1A - ON-SITE CONTAINED", "107. 8.1B - ON-SITE OTHER", "108. 8.1C - OFF-SITE CONTAIN", "109. 8.1D - OFF-SITE OTHER R", "110. 8.2 - ENERGY RECOVER ON", "111. 8.3 - ENERGY RECOVER OF", "112. 8.4 - RECYCLING ON SITE", "113. 8.5 - RECYCLING OFF SIT", "114. 8.6 - TREATMENT ON SITE", "115. 8.7 - TREATMENT OFF SITE", "117. 8.8 - ONE-TIME RELEASE")
tri_data_2011 <- read_csv("../data/2011_us.csv", show_col_types = FALSE)
tri_data_2011 <- tri_data_2011 %>% 
  select(my_variables)

```

```{r data-set-assignment-2012, warning = FALSE}
library(readr)
tri_data_2012 <- read_csv("../data/2012_us.csv", show_col_types = FALSE)
tri_data_2012 <- tri_data_2012 %>% 
  select(my_variables)
```

``` {r data-set-assignment-2013, warning = FALSE}
library(readr) 
tri_data_2013 <- read_csv("../data/2013_us.csv", show_col_types = FALSE)
tri_data_2013 <- tri_data_2013 %>% 
  select(my_variables)
```

``` {r data-set-assignment-2014, warning = FALSE}
library(readr)
tri_data_2014 <- read_csv("../data/2014_us.csv", show_col_types = FALSE)
tri_data_2014 <- tri_data_2014 %>% 
  select(my_variables)
```

```{r data-set-assignment-2015, warning = FALSE}
library(readr)
tri_data_2015 <- read_csv("../data/2015_us.csv", show_col_types = FALSE)
tri_data_2015 <- tri_data_2015 %>% 
  select(my_variables)
```

```{r data-set-assignment-2016, warning = FALSE}
library(readr)
tri_data_2016 <- read_csv("../data/2016_us.csv", show_col_types = FALSE)
tri_data_2016 <- tri_data_2016 %>% 
  select(my_variables)
```

```{r data-set-assignment-2017, warning = FALSE}
library(readr)
tri_data_2017 <- read_csv("../data/2017_us.csv", show_col_types = FALSE)
tri_data_2017 <- tri_data_2017 %>% 
  select(my_variables)
```

```{r data-set-assignment-2018, warning = FALSE}
library(readr)
tri_data_2018 <- read_csv("../data/2018_us.csv", show_col_types = FALSE)
tri_data_2018 <- tri_data_2018 %>% 
  select(my_variables)
```

```{r data-set-assignment-2019, warning = FALSE}
library(readr)
tri_data_2019 <- read_csv("../data/2019_us.csv", show_col_types = FALSE)
tri_data_2019 <- tri_data_2019 %>% 
  select(my_variables)
```

```{r data-set-assignment-2020, warning = FALSE}
library(readr)
tri_data_2020 <- read_csv("../data/2020_us.csv", show_col_types = FALSE)
tri_data_2020 <- tri_data_2020 %>% 
  select(my_variables)
```

```{r data-set-assignment-2021, warning = FALSE}
library(readr)
tri_data_2021 <- read_csv("../data/2021_us.csv", show_col_types = FALSE)
tri_data_2021 <- tri_data_2021 %>% 
  select(my_variables)
```

```{r bind-rows-2011-2021}
tri_data <- dplyr::bind_rows(tri_data_2011, tri_data_2012, tri_data_2013, tri_data_2014, tri_data_2015, tri_data_2016, tri_data_2017, tri_data_2018, tri_data_2019, tri_data_2020, tri_data_2021)
```

```{r recode-variables}
tri_data <- tri_data %>% 
  rename("year" = "1. YEAR", 
         "facility_name" = "4. FACILITY NAME",
         "street_address" = "5. STREET ADDRESS", 
         "city" = "6. CITY", 
         "county" = "7. COUNTY", 
         "state" = "8. ST", 
         "zip_code" = "9. ZIP", 
         "bia_code" = "10. BIA", 
         "tribal_land" = "11. TRIBE", 
         "latitude" = "12. LATITUDE", 
         "longitude" = "13. LONGITUDE", 
         "parent_co_name" = "15. PARENT CO NAME", 
         "standard_parent_co_name" = "17. STANDARD PARENT CO NAME", 
         "federal_facility" = "18. FEDERAL FACILITY", 
         "industry_sector" = "20. INDUSTRY SECTOR", 
         "chemical" = "34. CHEMICAL", 
         "elemental_metal_included" = "35. ELEMENTAL METAL INCLUDED", 
         "clean_air_act_chemical" = "39. CLEAN AIR ACT CHEMICAL", 
         "classification" = "40. CLASSIFICATION", 
         "metal" = "41. METAL", 
         "metal_category" = "42. METAL CATEGORY", 
         "carcinogen" = "43. CARCINOGEN", 
         "pbt" = "44. PBT", 
         "pfas" = "45. PFAS", 
         "form_type" = "46. FORM TYPE", 
         "unit_of_measurement" = "47. UNIT OF MEASURE", 
         "fugitive_air" = "48. 5.1 - FUGITIVE AIR", 
         "stack_air" = "49. 5.2 - STACK AIR", 
         "water" = "50. 5.3 - WATER", 
         "underground" = "51. 5.4 - UNDERGROUND", 
         "underground_class_1" = "52. 5.4.1 - UNDERGROUND CL I", 
         "underground_class_2_through_5" = "53. 5.4.2 - UNDERGROUND C II-V", 
         "landfills" = "54. 5.5.1 - LANDFILLS", 
         "land_treatment" = "57. 5.5.2 - LAND TREATMENT", 
         "surface_impoundment" = "58. 5.5.3 - SURFACE IMPNDMNT", 
         "other_disposal" = "61. 5.5.4 - OTHER DISPOSAL", 
         "on_site_release_total" = "62. ON-SITE RELEASE TOTAL", 
         "public_treatment_transfer_release" = "63. 6.1 - POTW - TRNS RLSE",
         "public_treatment_transfer_treat" = "64. 6.1 - POTW - TRNS TRT", 
         "public_treatment_total_transfer" = "65. POTW - TOTAL TRANSFERS", 
         "off_site_release_total" = "85. OFF-SITE RELEASE TOTAL", 
         "off_site_recycled_total" = "91. OFF-SITE RECYCLED TOTAL", 
         "off_site_energy_recovery_total" = "94. OFF-SITE ENERGY RECOVERY T", 
         "off_site_treated_total" = "101. OFF-SITE TREATED TOTAL", 
         "total_transfer" = "103. 6.2 - TOTAL TRANSFER", 
         "total_releases" = "104. TOTAL RELEASES", 
         "releases" = "105. 8.1 - RELEASES", 
         "on_site_contained" = "106. 8.1A - ON-SITE CONTAINED", 
         "on_site_other" = "107. 8.1B - ON-SITE OTHER", 
         "off_site_contain" = "108. 8.1C - OFF-SITE CONTAIN", 
         "off_site_other" = "109. 8.1D - OFF-SITE OTHER R", 
         "energy_recover_on" = "110. 8.2 - ENERGY RECOVER ON", 
         "energy_recover_off" = "111. 8.3 - ENERGY RECOVER OF", 
         "recycling_on_site" = "112. 8.4 - RECYCLING ON SITE", 
         "recycling_off_site" = "113. 8.5 - RECYCLING OFF SIT", 
         "treatment_on_site" = "114. 8.6 - TREATMENT ON SITE", 
         "treatment_off_site" = "115. 8.7 - TREATMENT OFF SITE", 
         "one_time_release" = "117. 8.8 - ONE-TIME RELEASE")
```

Preliminary data analysis is below. This data set includes 26,267 facilities, 6,345 parent companies, 30 industries, 627 chemicals, as well as a smaller dataset called chemical_info that holds information about each chemical and whether it is a metal, pbt, pfas, or carcinogen and if it is included in the Clean Air Act. 

```{r facilities}
tri_data %>% 
  select(facility_name) %>% 
  distinct(facility_name) %>% 
  arrange(facility_name)
```

```{r parent-companies}
tri_data %>% 
  select(parent_co_name) %>% 
  distinct(parent_co_name) %>% 
  arrange(parent_co_name)
```

```{r industry-sectors}
tri_data %>% 
  select(industry_sector) %>% 
  distinct(industry_sector) %>% 
  arrange(industry_sector)
```

```{r chemicals}
tri_data %>% 
  select(chemical) %>% 
  distinct(chemical) %>% 
  arrange(chemical)
```

```{r chemical-info-dataset}
chemical_info <- tri_data %>% 
  select(chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  arrange(chemical)
```


## 3. Data analysis plan
This is a very large dataset with a mix of character and numerical variables, so we anticipate that the first part of this project will be learning the best ways to visualize this data. As this is not a data set or subject that either of us have prior experience with, we will also be learning more about the data (including background vocabulary) during the first few weeks of working with it. 

We have consolidated data from 2011 to 2021 into one data frame we have called tri_data. We examined each of the variables and selected 58 that we believe will be helpful for our analysis. Through this project, we are hoping to learn more about methods of data visualization, especially when working with large, complex datasets. We are also hoping to learn more about data manipulation, communicating our findings, and how to share complex data in ways that are accurate, accessible, and engaging. 

We are specifically interested in mapping relationships between industry sectors, emissions, chemical information (carcinogens, pfas, metals, pbts, etc), and location. Questions that will be guiding us are: 
1) is there a correlation between industry sector and frequency of carcinogen, pfas, and pbt use?
2) is there a correlation between location and frequency of use of Clean Air Act chemicals?
3) how many facilities are located on tribal land and is there a significant difference in the type and amount of chemical release when compared to facilities not on tribal land?
4) what are the trends in pollution (chemical release) over the past 10 years?




