---
title: "Project Analysis [insert catchy name here]"
author: "Cedar and Joanna"
date: "`r Sys.Date()`"
output: html_document
---


#Library and Data Read-In

```{r load-packages, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(leaflet)
library(ggthemes)
library(ggridges)
library(ggmap)
library(lubridate)
library(stringr)
library(readr)
library(RColorBrewer)
library(forcats)
library(colorspace)
```

```{r load-data}
```


#Outline

##Introduction
introduction to dataset & summary stats
introduce primary research question: what is the relationship between industry sector and location, pollution, chemical use and has this shifted over the past 10 years?

```{r}
```


##Location Analysis
location of facilities, location trends, tribal lands, proximity
question: how many facilities are located on tribal land and is there a significant difference in the type and amount of chemical release when compared to facilities not on tribal land?

##Industry Analysis
industry sector trends (location, chemical use, pollution), changes in industry sector prevalence
question: is there a correlation between industry sector and frequency of carcinogen, pfas, and pbt use?

##Chemical Analysis
chemical info (type, category, etc), trends in chemical use and disposal 
question: is there a correlation between location and frequency of use of Clean Air Act chemicals?

##Pollution analysis
trends in pollution, on-site vs off-site disposal
question: what are the trends in pollution (chemical release) over the past 10 years?

##Conclusion
overview of data and conclusions about pollution and chemical use trends