---
title: "Project Analysis [insert catchy name here]"
author: "Cedar and Joanna"
date: "`r Sys.Date()`"
output: html_document
---


#Library and Data Read-In

```{r load-packages, echo = FALSE, message = FALSE, warning = FALSE}
library(sf)
library(htmltools)
library(leafsync)
library(rgeoboundaries)
library(kableExtra)
library(tidyverse)
library(leaflet)
library(ggthemes)
library(ggridges)
library(ggmap)
library(lubridate)
library(stringr)
library(readr)
library(broom)
library(RColorBrewer)
library(forcats)
library(colorspace)
```

```{r data-read-in}
tri_data <- read_csv("../data/tri_data.csv", show_col_types = FALSE)
```


#Outline

##Introduction
introduction to dataset & summary stats
introduce primary research question: what is the relationship between industry sector and location, pollution, chemical use and has this shifted over the past 10 years?

##Location Analysis
location of facilities, location trends, tribal lands, proximity
question: how many facilities are located on tribal land and is there a significant difference in the type and amount of chemical release when compared to facilities not on tribal land?

```{r facility-location-dataset}
facility_location <- tri_data %>% 
  select("facility_name", "street_address", "city", "county", "state", "zip_code", "bia_code", "tribal_land", "latitude", "longitude") %>% 
  distinct()
```


```{r facility-map}
facility_location %>% 
  leaflet() %>% 
addTiles() %>% 
  addCircleMarkers(lng = ~longitude, 
                 lat = ~latitude, 
                 clusterOptions = markerClusterOptions())
```


##Industry Analysis
industry sector trends (location, chemical use, pollution), changes in industry sector prevalence
question: is there a correlation between industry sector and frequency of carcinogen, pfas, and pbt use?

```{r industry-sector-dataset}
industry_info <- tri_data %>% 
  select(industry_sector, state, clean_air_act_chemical, carcinogen, metal_category, pbt, pfas, on_site_release_total, off_site_release_total, on_site_contained, off_site_contain) %>% 
  distinct(industry_sector, state, clean_air_act_chemical, carcinogen, metal_category, pbt, pfas, on_site_release_total, off_site_release_total, on_site_contained, off_site_contain) %>% 
```


##Chemical Analysis
chemical info (type, category, etc), trends in chemical use and disposal 
question: is there a correlation between location and frequency of use of Clean Air Act chemicals?

```{r chemical-info-dataset}
chemical_info <- tri_data %>% 
  select(chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  arrange(chemical)
```


##Pollution analysis
trends in pollution, on-site vs off-site disposal
question: what are the trends in pollution (chemical release) over the past 10 years?

##Conclusion
overview of data and conclusions about pollution and chemical use trends