---
title: "Project Analysis [insert catchy name here]"
author: "Cedar Callahan and Joanna Pittari"
date: "`r Sys.Date()`"
output: html_document
---

# Library and Data Read-In

```{r load-packages, echo = FALSE, message = FALSE, warning = FALSE}
library(sf)
library(htmltools)
library(leafsync)
library(kableExtra)
library(tidyverse)
library(leaflet)
library(ggthemes)
library(ggridges)
library(ggmap)
library(lubridate)
library(stringr)
library(readr)
library(broom)
library(RColorBrewer)
library(forcats)
library(colorspace)
library(visdat)
library(naniar)
library(ggrepel)
library(styler)
library(lintr)
```

```{r data-read-in, echo = FALSE, message = FALSE, warning = FALSE}
tri_data <- read_csv("../data/tri_data.csv", show_col_types = FALSE)
```

# Introduction

The Environmental Protection Agency's database of Toxic Release Inventory lacks consistency in its observations and variables -- though this is due to the unexplained environment that informs the data. 

Every year, new chemicals to be reported to the EPA are added to the TRI's list, which means the amount of chemical variables within this one decade of 2011-2021 goes from 513 reported chemicals to 545 reported chemicals which actually accounts for around 200 additions and many undocumented deletions of variables (not counting the 16 that were added to the dataset in 2011). Rules on what qualifies as the threshold for reporting also change as certain chemicals fall in and out of especially-close watch. Countrywide and statewide laws on environmental protection grant even more opportunity for data to drastically change year by year, though in technicality, the actual happenings at these facilities may not be changing much at all. Some of these laws -- though not implying causation -- will be clarified where relevant in the code.

Our goal is to use what data we find (that is reasonably accurate given the circumstances) in order to identify the layout of these waste chemicals across the United States. To reach this, we must ask some guiding questions: What are the relationships between differing industry sectors and their growing or declining locations? Are there any shifts in the levels of overall pollution during the past decade, and are these changes set in any specific part of the country? Finally, how has chemical use and waste (as well as waste procedures) changed in the past decade, and what do we know about these chemicals from the EPA to infer their threat to human health?

# Data Overview

## Location Analysis

Our first question of locations deals with the trends of facilities' presence, whether or not they are on tribal land, and their proximity to each other (possibly by industry sector) -- which may help in assessing what is in the best interest of the people due to identifying the patterns of popularity at the same time as keeping safety and land protection a priority.

```{r facility-location-dataset, echo = FALSE, message = FALSE, warning = FALSE}
facility_location <- tri_data %>% 
  select("facility_name", "street_address", "city", "county", "state", "zip_code", "bia_code", "tribal_land", "latitude", "longitude") %>% 
  distinct()

facility_location <- facility_location %>% 
  mutate(tribal_yes_no = case_when(is.na(tribal_land) == FALSE ~ "Tribal Land",
                                   TRUE ~ "Not Tribal Land"))

tri_data <- tri_data %>% 
  mutate(tribal_yes_no = case_when(is.na(tribal_land) == FALSE ~ "Tribal Land",
                                   TRUE ~ "Not Tribal Land"))
```

### Number of Facilities Owned by Each Company

The five most common parent companies are as follows: U.S. Department of Defense, CRH Americas Inc, Cemex Inc, Argos U.S. Corp, and Berkshire Hathaway Inc. The first listed is a government agency responsible for U.S. military protection, followed by four different building/construction/transport materials companies.

```{r number-of-facilities, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  select(standard_parent_co_name, parent_co_name, facility_name) %>% 
  group_by(standard_parent_co_name, parent_co_name) %>% 
  summarise(number_of_facilities = n_distinct(facility_name)) %>% 
  arrange(desc(number_of_facilities))
```

### Map of Facilities

The large majority of facilities in this dataset are located in North America, but a significant number of U.S.-owned facilities are strewn across the global map in other territories -- possibly shifting any inferences that can be made on the state of pollution solely in the U.S.

```{r facility-map, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  distinct(longitude, latitude, facility_name, parent_co_name, industry_sector, city, state) %>% 
  leaflet() %>% 
  addTiles() %>% 
    addCircleMarkers(lng = ~longitude, 
                 lat = ~latitude, 
                 clusterOptions = markerClusterOptions(),
                 popup = ~paste0(facility_name, "<br>", parent_co_name, "<br>", industry_sector, "<br>", city, ", ", state))
```

### Percent of Facilities on Tribal Land

Though the percentage of facilities located on tribal land is less than 1% (there are `sum(facility_location$tribal_yes_no == "YES")` facilities on tribal land), and we are curious about any differences in the amount of chemicals used and released on tribal vs not tribal land.

```{r percent-tribal-land, echo = FALSE, message = FALSE, warning = FALSE}
facility_location %>% 
  select(tribal_yes_no) %>% 
  group_by(tribal_yes_no) %>%
  summarise(percent = 100 * n()/nrow(facility_location))

facility_location %>% 
  select(tribal_yes_no) %>% 
  group_by(tribal_yes_no) %>% 
  count(tribal_yes_no)
```

Using these tables, we have set up a groundwork map for placing chemical usage, which leads us to ask: knowing the drastic difference in percent of facilities on tribal land or not, does the location of tribal land have any effect on the type and amount of chemical release when compared to facilities not on tribal land?

We can take points from history and U.S. law, knowing that tribes of the American West often do have significant health hazards due to the many power plants, factories, mines, waste dumps, and testing sites that are placed on their land -- sometimes with permission as a means to acquiring much-needed funds, and sometimes out of a tribe's power. This question will appear further in visualizations.

## Industry Analysis

Within these observations of location, there is the question of each facility's purpose. Condensing the purpose of a factory into just 11 categories, their individual patterns of location, chemical use, and pollution can be tracked. This prevalence allows us to ask: is there a correlation between an industry's sector and their frequency of chemicals? Further, their frequency of specific types of chemicals -- carcinogens, PFAS', and PBT'? 

```{r industry-sector-dataset, echo = FALSE, message = FALSE, warning = FALSE}
industry_info <- tri_data %>% 
  select(industry_sector, state, clean_air_act_chemical, carcinogen, metal_category, pbt, pfas, on_site_release_total, off_site_release_total, on_site_contained, off_site_contain) %>% 
  distinct(industry_sector, state, clean_air_act_chemical, carcinogen, metal_category, pbt, pfas, on_site_release_total, off_site_release_total, on_site_contained, off_site_contain)
```

```{r industry-sector-categories, echo = FALSE, message = FALSE, warning = FALSE}
tri_data <- tri_data %>% 
  mutate(industry_sector_category = case_when(industry_sector %in% c("Tobacco", "Beverages", "Food") ~ "Agriculture and Food Processing",
  industry_sector %in% c("Electrical Equipment", "Transportation Equipment", "Machinery", "Computers and Electronic     Products") ~ "Machinery and Technology",
  industry_sector %in% c("Miscellaneous Manufacturing", "Other") ~ "Misc",
  industry_sector %in% c("Chemicals", "Chemical Wholesalers", "Hazardous Waste") ~ "Chemicals and Waste",
  industry_sector %in% c("Leather", "Apparel", "Textiles", "Textile Product") ~ "Textiles and Leather",
  industry_sector %in% c("Petroleum Bulk Terminals", "Petroleum") ~ "Petroleum",
  industry_sector %in% c("Metal Mining", "Coal Mining") ~ "Mining",
  industry_sector %in% c("Nonmetallic Mineral Product", "Plastics and Rubber", "Fabricated Metals", "Primary Metals") ~ "Processed Materials",
  industry_sector %in% c("Furniture", "Wood Products", "Paper", "Printing", "Publishing") ~ "Wood and Paper",
  industry_sector %in% "Electric Utilities" ~ "Electric Utilities"))
```

### Industry Sector Distribution

The most common purpose of a U.S. facility is dealing with the production, waste, and wholesales of chemicals, accounting for almost 30% of U.S. industries as shown in the table. The least common accounting for about 0.4% is textile production which makes sense given trends in American imports.

```{r industry-sector-category-distribution, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
 select(industry_sector_category) %>% 
 group_by(industry_sector_category) %>%
 summarise(percent = 100 * n()/nrow(tri_data)) %>% 
 arrange(desc(percent))

tri_data %>% 
  ggplot(aes(y = fct_infreq(industry_sector_category), fill = industry_sector_category)) +
  geom_bar() +
  labs(title = "Industry Sector Prevalence",
    subtitle = "Distribution of Industry Sector Categories in the TRI",
    x = "Count",
    y = "Industry Sector Category",
    fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d()
```

### Carcinogen Percentage by Industry Sector Graph

Understandably fitting with the most common industry sector, the chemical-producing facilities also use the most carcinogens (non-comprehensive), but the frequency seems to [almost] consistently grow with the prevalence of these sectors. That being said, every sector listed has a chemical usage that is majority non-carcinogenic.

```{r industry-sector-and-carcinogen-use, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  ggplot(aes(y = fct_infreq(industry_sector), fill = carcinogen)) +
  geom_bar() +
  labs(title = "Carcinogen Use by Industry Sector",
       x = "Count",
       y = "Industry Sector",
       fill = "Carcinogenic") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)
```

## Chemical Analysis

Honing our observational questions down to just chemical trends of use and disposal now, we see how the laws of the EPA inform chemical procedures. Specifically in terms of the Clean Air Act, is there a differing correlation between the location and frequency of these specially-listed chemicals when compared to chemicals not listed on the Clean Air Act? And more broadly, is there any type, category, or closely-watched category of chemical that is most common despite the efforts of these laws?

```{r chemical-info-dataset, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info <- tri_data %>% 
  select(year, chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  arrange(chemical)
```

The most commonly used chemical is lead, accounting for over 5% of use, followed by its compounds that make up almost 4% of use. Lead is known to cause poisoning of the human body, leading to neurological and reproductive health issues as well as impaired function of the kidneys and other body systems. Poisoning can be acquired through contaminated dust particles and products, contaminated drinking water, and contaminated soil.

```{r chemical-percent, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical) %>% 
  group_by(chemical) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))
```

### Chemical Category Calculations

The large majority of chemicals used are not elemental metals or metals in general. About 63% are identified in the Clean Air Act, leaving 37% not present. Over 28% of chemicals used in these facilities are known to be carcinogenic, 17% are PBTs, and 0.02% are PFAS'. These smaller percentages are still a concern, and are in-part new variables to track. PFAS did not have to be reported to the TRI until 2020 (making this decade of data insufficient), even though over 2,000 facilities were known to be producing them before the regulation. Therefore, it has far fewer observations in comparison to other chemicals though used since the 1940s and known to cause certain cancers, impairment of the immune system, reproductive issues, and abnormal developmental effects. The reporting threshold for that particular subset of chemicals is 100 pounds per year.

```{r chemical-percent-elemental-metal-included, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(elemental_metal_included) %>% 
  group_by(elemental_metal_included) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>%
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = elemental_metal_included, fill = elemental_metal_included)) +
  geom_bar() +
  labs(title = "Elemental Metal Prevalence by Year",
       x = "Elemental Metal",
       y = "Count",
       fill = "Elemental Metal") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-percent-clean-air-act, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(clean_air_act_chemical) %>% 
  group_by(clean_air_act_chemical) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = clean_air_act_chemical, fill = clean_air_act_chemical)) +
  geom_bar() +
  labs(title = "Clean Air Act Chemical Prevalence by Year",
       x = "Clean Air Act Chemical", 
       y = "Count",
       fill = "Clean Air Act Chemical") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year)
```

```{r chemical-percent-metal, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(metal) %>% 
  group_by(metal) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = metal, fill = metal)) +
  geom_bar() +
  labs(title = "Metal Prevalence by Year",
       x = "Metal",
       y = "Count",
       fill = "Metal") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-percent-metal-category, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(metal_category) %>% 
  group_by(metal_category) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = metal_category, fill = metal_category)) +
  geom_bar() +
  labs(title = "Metal Category Prevalence by Year",
       x = "Metal Category",
       y = "Count",
       fill = "Metal Category") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~year) 

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = carcinogen, fill = carcinogen)) +
  geom_bar() +
  labs(title = "Carcinogen Prevalence by Year",
       x = "Carcinogenic",
       y = "Count",
       fill = "Carcinogenic") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-trends-by-category-pbt, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = pbt, fill = pbt)) +
  geom_bar() +
  labs(title = "PBT Prevalence by Year",
       x = "PBT",
       y = "Count",
       fill = "PBT") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-trends-by-category-pfas, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = pfas, fill = pfas)) +
  geom_bar() +
  labs(title = "PFAS Prevalence by Year",
       x = "PFAS",
       y = "Count",
       fill = "PFAS") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

For above visualizations: elemental metals (some harmless, some causing neurotoxicity) keep a near-consistent distribution in the chemical data. The few increases this type of metal makes in the data is adjusted by the many additions of tracked chemicals in the dataframe, also causing the count for non-elemental-metals to increase. A similar state of increase goes for the amount of Clean Air Act chemicals and their growing amendments (still with more Clean Air Act chemicals than not) and with the general metals graph (more non-metals than metals, though both counts increasing). The amount of PBTs remains consistent with the non-PBTs increasing with new variables, and our two years of PFAS data show their consistency -- though, of course, not much judgement can be made through this small piece of information, and the same must go for much of these environmentally-altered graphs.

##### Metal Category Pie Chart

```{r metal-category-pie-chart, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  group_by(metal_category) %>% 
  summarise(prop = n()/nrow(tri_data)) %>% 
  ggplot(aes(x = "", y = prop, fill = fct_inorder(metal_category))) +
  geom_col(color = "white", linewidth = .3) +
  geom_text(aes(x = "", y = prop, label = round(prop, 2)), position = position_stack(vjust = 0.5)) +
  labs(title = "Metal Category Distribution",
       x = "",
       y = "", 
       fill = "Metal Category") +
  theme_void() +
  coord_polar(theta = "y") +
  scale_fill_viridis_d(begin = 0.4, end = 0.9)
```

```{r chemical-percent-carcinogen, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(carcinogen) %>% 
  group_by(carcinogen) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = carcinogen, fill = carcinogen)) +
  geom_bar() +
  labs(title = "Carcinogen Prevalence by Year",
       x = "Carcinogenic",
       y = "Count",
       fill = "Carcinogenic") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year)
```

```{r chemical-percent-pbt, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(pbt) %>% 
  group_by(pbt) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = pbt, fill = pbt)) +
  geom_bar() +
  labs(title = "PBT Prevalence by Year",
       x = "PBT",
       y = "Count",
       fill = "PBT") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-percent-pfas, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(pfas) %>% 
  group_by(pfas) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = pfas, fill = pfas)) +
  geom_bar() +
  labs(title = "PFAS Prevalence by Year",
       x = "PFAS",
       y = "Count",
       fill = "PFAS") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

For above visualizations: elemental metals (some harmless, some causing neurotoxicity) keep a near-consistent distribution in the chemical data. The few increases this type of metal makes in the data is adjusted by the many additions of tracked chemicals in the dataframe, also causing the count for non-elemental-metals to increase. A similar state of increase goes for the amount of Clean Air Act chemicals and their growing amendments (still with more Clean Air Act chemicals than not) and with the general metals graph (more non-metals than metals, though both counts increasing). The amount of PBTs remains consistent with the non-PBTs increasing with new variables, and our two years of PFAS data show their consistency -- though, of course, not much judgement can be made through this small piece of information, and the same must go for much of these environmentally-altered graphs.


### Chemical Categories

There are `sum(chemical_info$carcinogen == "YES")` carcinogens being observed -- a large number creating a small percentage.

```{r carcinogen-percent-yes, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, carcinogen) %>% 
  filter(carcinogen == "YES") %>% 
  group_by(chemical)
```

There are `sum(chemical_info$carcinogen == "NO")` non-carcinogens being observed.

```{r carcinogen-percent-no, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, carcinogen) %>% 
  filter(carcinogen == "NO") %>% 
  group_by(chemical)
```

There are `sum(chemical_info$pfas == "YES")` observations of PFAS chemicals in the dataset.

```{r pfas-percent-yes, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pfas) %>% 
  filter(pfas == "YES") %>% 
  group_by(chemical)
```

There are `sum(chemical_info$pfas == "NO")` observations of non-PFAS chemicals in the dataset.

```{r pfas-percent-no, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pfas) %>% 
  filter(pfas == "NO") %>% 
  group_by(chemical)
```

There are `sum(chemical_info$pbt == "YES")` observations of PBT chemicals in the dataset.

```{r pbt-percent-yes, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pbt) %>% 
  filter(pbt == "YES") %>% 
  group_by(chemical)
```

There are `sum(chemical_info$pbt == "NO")` observations of non-PBT chemicals in the dataset.

```{r pbt-percent-no, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pbt) %>% 
  filter(pbt == "NO") %>% 
  group_by(chemical)
```

## Pollution analysis

From these distributions of observed chemicals, there is still the question of their disposal after use. With air, ground, and water being some possible points of disposal, we must assess the trends in these differing types of waste removal. In the past decade, what are these trends in pollution due to chemical release, whether on-site or off?

```{r pollution-info-dataset, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info <- tri_data %>% 
  select(year, fugitive_air, stack_air, water, underground, underground_class_1, underground_class_2_through_5, landfills, land_treatment, surface_impoundment, other_disposal, on_site_release_total, public_treatment_total_transfer, off_site_release_total, off_site_recycled_total, off_site_energy_recovery_total, off_site_treated_total, total_transfer, total_releases, releases, on_site_contained, on_site_other, off_site_contain, off_site_other)
```

### On-Site Pollution by Category

#### By Type

On-site release types: variables `Underground`, `Landfill`, and `Surface Impoundment` are all zero.

```{r pollution-by-on-site-release-fugitive-air, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = fugitive_air, color = fugitive_air, fill = fugitive_air)) + 
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Fugitive Air Release by Year",
       x = "On Site Release Total",
       y = "Fugitive Air") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-stack-air, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = stack_air)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Stack Air Release by Year",
       x = "On Site Release Total",
       y = "Stack Air") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-water, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = water)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Water Release by Year",
       x = "On Site Release Total",
       y = "Water") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-underground-class-1, echo = FALSE, message = FALSE, warning = FALSE}
 pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_1)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Underground Class 1 Releases by Year",
       x = "On Site Release Total",
       y = "Underground Class 1") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-underground-class-2-to-5, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_2_through_5)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Underground Class 2 Through 5 Release by Year",
       x = "On Site Release Total",
       y = "Underground Class 2 Through 5") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-land-treatment, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = land_treatment)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Land Treatment by Year",
       x = "On Site Release Total",
       y = "Land Treatment") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-other-disposal, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = other_disposal)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Other Disposal by Year",
       x = "On Site Release Total",
       y = "Other Disposal") +
  theme_minimal() +
  facet_wrap(~ year)
```

#### By Year

```{r on-site-release-total-by-year, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = on_site_release_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "On Site Release Total from 2011 to 2021",
       x = "Year",
       y = "On Site Release Total") +
  theme_minimal() 
```

### Pollution by General Category (Including On-site-total)

Pollution variable `Releases` is all zeros.

```{r pollution-by-general-category, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = on_site_release_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "On Site Release Total by Year",
       x = "Year",
       y = "On Site Release Total") +
  theme_minimal()
```

```{r pollution-public-treatment-total-transfer, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = public_treatment_total_transfer)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Transfer to Public Treatment by Year",
       x = "Year",
       y = "Total Transfer to Public Treatment") +
  theme_minimal()
```

```{r pollution-off-site-release-total, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_release_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Off Site Release Total by Year",
       x = "Year",
       y = "Off Site Release Total") +
  theme_minimal()
```

```{r pollution-off-site-recycled-total, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_recycled_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Recycled Off Site by Year",
       x = "year",
       y = "Total Recycled Off Site") +
  theme_minimal()
```

```{r pollution-off-site-energy-recovery-total, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_energy_recovery_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Energy Recovery Off Site by Year",
       x = "Year",
       y = "Total Energy Recovery Off Site") +
  theme_minimal()
```

```{r pollution-off-site-treated-total, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_treated_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Treated Off Site by Year",
       x = "Year",
       y = "Total Treated Off Site") +
  theme_minimal()
```

```{r pollution-total-transfer, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = total_transfer)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Transfered by Year",
       x = "Year",
       y = "Total Transfered") +
  theme_minimal()
```

```{r pollution-total-releases, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = total_releases)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Released by Year",
       x = "Year", 
       y = "Total Released")
```

```{r pollution-on-site-contained, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = on_site_contained)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Amount Contained On Site By Year",
       x = "Year",
       y = "Amount Contained On Site") +
  theme_minimal()
```

```{r pollution-on-site-other, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = on_site_other)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "On Site 'Other' by Year",
       x = "Year",
       y = "On Site 'Other'") +
  theme_minimal()
```

```{r pollution-off-site-contain, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_contain)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Contained Off Site by Year",
       x = "Year",
       y = "Contained Off Site") +
  theme_minimal()
```

```{r pollution-off-site-other, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_other)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Off Site 'Other' by Year",
       x = "Year",
       y = "Off Site 'Other'") +
  theme_minimal()
```



```{r pollution-water-log-graph, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>%
  filter(water > 0) %>%
  ggplot(aes(x = log10(on_site_release_total), y = water/on_site_release_total, color = industry_sector_category)) +
  geom_point(alpha = 0.3) +
  labs(title = "Water Release by Industry Sector",
       subtitle = "Log10 Analysis of Water Release Divided by On Site Release Total",
       x = "On Site Release Total",
       y = "Water Release",
       color = "") +
  theme_minimal() +
  scale_color_viridis_d() +
  facet_wrap(~ industry_sector_category)
```


```{r industry-sector-category-water-release, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  ggplot(aes(y = industry_sector_category, x = water, fill = industry_sector_category)) +
  geom_col() +
  labs(title = "Water Release by Industry Sector Category",
       x = "Water",
       y = "Industry Sector Category",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d()
```

# Detailed Analysis

```{r create-analysis-data-frame, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data <- tri_data %>% 
  filter(chemical %in% c("Arsenic", "Styrene", "Formaldehyde", "Nickel", "Chromium", "Lead", "Mercury", "Ammonia", "Phosphorus (yellow or white)", "Chloroform", "Benzo[g,h,i]perylene")) 

chemical_analysis_data <- chemical_analysis_data %>%
  mutate(off_site_total = off_site_release_total + off_site_recycled_total + off_site_energy_recovery_total + public_treatment_total_transfer + off_site_treated_total + total_transfer + off_site_contain + off_site_other)

chemical_analysis_data <- chemical_analysis_data %>% 
  mutate(on_site_total = fugitive_air + stack_air + water + underground + underground_class_1 + underground_class_2_through_5 + landfills + land_treatment + surface_impoundment + other_disposal + on_site_release_total + on_site_contained + on_site_other)
```

## Chemicals and Location

```{r count-chemical-and-type, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  count(chemical) %>% 
  arrange(desc(n))

chemical_analysis_data %>% 
  select(chemical, pfas, pbt, clean_air_act_chemical, carcinogen) %>% 
  group_by(chemical) %>% 
  count(pfas, pbt, clean_air_act_chemical, carcinogen) %>% 
  arrange(desc(n))
```

```{r chemical-use-by-facility, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  select(standard_parent_co_name, parent_co_name, facility_name, chemical) %>% 
  group_by(standard_parent_co_name, parent_co_name, facility_name) %>% 
  summarise(number_of_chemicals = n_distinct(chemical)) %>% 
  arrange(desc(number_of_chemicals))
```

```{r chemical-use-on-tribal-lands, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  group_by(tribal_land, chemical) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = chemical, y = prop, fill = tribal_land)) +
  geom_col(position = "fill") +
  labs(title = "Chemical Use On and Off Tribal Lands",
       subtitle = "By Chemical",
       x = "Chemical",
       y = "Proportion") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  select(chemical, tribal_land) %>% 
  count(chemical, tribal_land)

chemical_analysis_data %>% 
  ggplot(aes(y = tribal_land, fill = chemical)) +
  geom_bar(position = "fill", color = "white") +
  labs(title = "Chemical Use On and Off Tribal Lands",
       subtitle = "By Tribal and Not Tribal Land",
       y = "Tribal Land?") +
  theme_minimal() +
  scale_fill_viridis_d()
```

```{r chemical-use-by-state,echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>%
  select(state, chemical) %>%
  group_by(state) %>%
  summarise(chemical_use = n_distinct(chemical)) %>%
  arrange(desc(chemical_use))
```

### Overview graphs

## Chemical Release Graphs

```{r chemical-category-graphs, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = pbt)) +
  geom_bar() +
  labs(title = "Is This Chemical a PBT?",
       y = "Chemical",
       fill = "PBT") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = pfas)) +
  geom_bar() +
  labs(title = "Is This Chemical a PFAS?",
       y = "Chemical",
       fill = "PFAS") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = carcinogen)) +
  geom_bar() +
  labs(title = "Is This Chemical Carcinogenic?",
       y = "Chemical",
       fill = "Carcinogenic") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = clean_air_act_chemical)) +
  geom_bar() +
  labs(title = "Is This Chemical Included in the Clean Air Act?",
       y = "Chemical",
       fill = "Clean Air Act Chemical?") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)
```

### Graphs of on-site disposal by type

### Graphs of on-site release by type

```{r plot-fugitive-air-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = fugitive_air)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Fugitive Air in On Site Release Total",
       x = "On Site Release Total",
       y = "Fugitive Air",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-stack-air-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = stack_air)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Stack Air in On Site Release Total",
       x = "On Site Release Total",
       y = "Stack Air",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-water-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = water)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Water in On Site Release Total",
       x = "On Site Release Total",
       y = "Water",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-underground-class-1-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_1)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Underground Class 1 in On Site Release Total",
       x = "On Site Release Total",
       y = "Underground Class 1",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-underground-2-to-5-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_2_through_5)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Underground Class 2 through 5 in On Site Release Total",
       x = "On Site Release Total",
       y = "Underground Class 2 through 5",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-land-treatment-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = land_treatment)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Land Treatment in On Site Release Total",
       x = "On Site Release Total",
       y = "Land Treatment",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-other-disposal-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = other_disposal)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "'Other' Disposal in On Site Release Total",
       x = "On Site Release Total",
       y = "'Other' Disposal",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-on-site-release-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = chemical, fill = chemical)) +
  geom_col() +
  labs(title = "On Site Release Total by Chemical",
       x = "On Site Release Total",
       y = "Chemical",
       fill = "Chemical") +
  theme_minimal() +
  scale_fill_viridis_d()
```

### Graphs of on-site disposal by type

```{r plot-on-site-contained-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_total, y = on_site_contained)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Contained On Site by Chemical",
       x = "On Site Total",
       y = "On Site Contained",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-on-site-other-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_total, y = on_site_other)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "On Site 'Other' and On Site Total",
       x = "On Site Total",
       y = "On Site 'Other'",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

### Graphs of off-site disposal by type

```{r plot-public-treatment-transfer-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = public_treatment_total_transfer)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Transfered to Public Treatment and Off Site Total",
       x = "Off Site Total",
       y = "Total Transfered to Public Treatment",
       color = "Chemical",
       alpha = "Saturation")  +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-release-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_release_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Off Site Release Total and Off Site Total",
       x = "Off Site Total",
       y = "Off Site Release Total",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-recycled-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_recycled_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Recycled off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Recycled Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-energy-recovery-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_energy_recovery_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Energy Recovered Off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Energy Recovered Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-treated-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_treated_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Treated Off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Treated Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-total-transfer-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = total_transfer)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Transfered and Off Site Total",
       x = "Off Site Total",
       y = "Total Transfered",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-total-releases-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = total_releases)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Released and Off Site Total",
       x = "Off Site Total",
       y = "Total Released",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-contain-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_contain)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Contained Off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Contained Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-other-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_other)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Off Site 'Other' and Off Site Total",
       x = "Off Site Total",
       y = "Off Site 'Other'",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```


### Chemical Release Trends by Year

Note: Phosphorus is not faceted by industry sector because there is only one industry that uses it.

```{r on-site-total-by-year}
chemical_analysis_data %>% 
  ggplot(aes(x = chemical, y = on_site_total, fill = chemical)) +
  geom_col(position = "dodge") +
  labs(title = "Total On Site By Chemical",
       subtitle = "On Site Release, Storage, and Disposal",
       x = "Chemical",
       y = "On Site Total",
       fill = "Chemical") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~ year)
```

```{r off-site-total-by-year}
chemical_analysis_data %>% 
  ggplot(aes(x = chemical, y = off_site_total, fill = chemical)) +
  geom_col(position = "dodge") +
    labs(title = "Total Off Site By Chemical",
       subtitle = "Off Site Release, Storage, Treatment, and Disposal",
       x = "Chemical",
       y = "Off Site Total",
       fill = "Chemical") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~ year)
```

```{r arsenic-release-by-year}
chemical_analysis_data %>%
  filter(chemical == "Arsenic") %>%
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Arsenic Off Site Total",
       subtitle = "Arsenic Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>%
  filter(chemical == "Arsenic") %>%
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Arsenic On Site Total",
       subtitle = "Arsenic On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r styrene-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Styrene") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Styrene Off Site Total",
       subtitle = "Styrene Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Styrene") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Styrene On Site Total",
       subtitle = "Styrene On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r formaldehyde-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Formaldehyde") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Formaldehyde Off Site Total",
       subtitle = "Formaldehyde Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Formaldehyde") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Formaldehyde On Site Total",
       subtitle = "Formaldhyde On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r nickel-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Nickel") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Nickel Off Site Total",
       subtitle = "Nickel Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Nickel") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Nickel On Site Total",
       subtitle = "Nickel On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +  
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r chromium-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Chromium") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Chromium Off Site Total",
       subtitle = "Chromium Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Chromium") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Nickel On Site Total",
       subtitle = "Nickel On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r lead-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Lead") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Lead Off Site Total",
       subtitle = "Lead Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Lead") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Lead On Site Total",
       subtitle = "Lead On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r mercury-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Mercury") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Mercury Off Site Total",
       subtitle = "Mercury Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Mercury") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Mercury On Site Total",
       subtitle = "Mercury On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r ammonia-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Ammonia") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Ammonia Off Site Total",
       subtitle = "Ammonia Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Ammonia") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Ammonia On Site Total",
       subtitle = "Ammonia On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r phosphorus-release-by-year}
chemical_analysis_data %>%
  filter(chemical == "Phosphorus") %>%
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Phosphorus Off Site Total",
       subtitle = "Phosphorus Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
  fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d()

chemical_analysis_data %>%
  filter(chemical == "Phosphorus") %>%
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Phosphorus On Site Total",
      subtitle = "Phosphorus On Site Release, Storage, and Disposal",
      x = "Year",
      y = "On Site Total",
      fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d()
```

```{r chloroform-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Chloroform") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Chloroform Off Site Total",
       subtitle = "Chloroform Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Chloroform") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Chloroform On Site Total",
       subtitle = "Chloroform On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r benzo[g,h,i]perylene-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Benzo[g,h,i]perylene") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Benzo[g,h,i]perylene Off Site Total",
       subtitle = "Benzo[g,h,i]perylene Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Benzo[g,h,i]perylene") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Benzo[g,h,i]perylene On Site Total",
       subtitle = "Benzo[g,h,i]perylene On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```


# Conclusion

In recognizing the increase of observations, increase in pollution trends, and the environment of increasingly strict reporting laws, we must be careful not to make the assumption that pollution is truly on the rise, and that conditions are getting worse in general. What the legal environment shows is an increase in control and recognition of needed chemical regulations. What the work shows is a very actively polluted environment irregardless of these laws, and the continued prevalence of chemicals that are known to have been used as weapons of war and for body-preservation and alteration -- clearly harmful to human health. This is not to say that there is no improvement, but that there surely is much more that must be done.