---
title: "Project Analysis [insert catchy name here]"
author: "Cedar Callahan and Joanna Pittari"
date: "`r Sys.Date()`"
output: html_document
---

# Library and Data Read-In

```{r load-packages, echo = FALSE, message = FALSE, warning = FALSE}
library(sf)
library(htmltools)
library(leafsync)
library(kableExtra)
library(tidyverse)
library(leaflet)
library(ggthemes)
library(ggridges)
library(ggmap)
library(lubridate)
library(stringr)
library(readr)
library(broom)
library(RColorBrewer)
library(forcats)
library(colorspace)
library(visdat)
library(naniar)
library(ggrepel)
library(styler)
library(lintr)
```

```{r data-read-in, echo = FALSE, message = FALSE, warning = FALSE}
tri_data <- read_csv("../data/tri_data.csv", show_col_types = FALSE)
```

# Introduction

The Environmental Protection Agency's database of Toxic Release Inventory lacks consistency in its observations and variables -- though this is due to the unexplained environment that informs the data.

Every year, new chemicals to be reported to the EPA are added to the TRI's list, which means the amount of chemical variables within this one decade of 2011-2021 goes from 513 reported chemicals to 545 reported chemicals which actually accounts for around 200 additions and many undocumented deletions of variables (not counting the 16 that were added to the dataset in 2011). Rules on what qualifies as the threshold for reporting also change as certain chemicals fall in and out of especially-close watch. Countrywide and statewide laws on environmental protection grant even more opportunity for data to drastically change year by year, though in technicality, the actual happenings at these facilities may not be changing much at all. Some of these laws -- though not implying causation -- will be clarified where relevant in the code.

Our goal is to use what data we find (that is reasonably accurate given the circumstances) in order to identify the layout of these waste chemicals across the United States. To reach this, we must ask some guiding questions: What are the relationships between differing industry sectors and their growing or declining locations? Are there any shifts in the levels of overall pollution during the past decade, and are these changes set in any specific part of the country? Finally, how has chemical use and waste (as well as waste procedures) changed in the past decade, and what do we know about these chemicals from the EPA to infer their threat to human health?

# Data Overview

## Location Analysis

Our first question of locations deals with the trends of facilities' presence, whether or not they are on tribal land, and their proximity to each other (possibly by industry sector) -- which may help in assessing what is in the best interest of the people due to identifying the patterns of popularity at the same time as keeping safety and land protection a priority.

```{r facility-location-dataset, echo = FALSE, message = FALSE, warning = FALSE}
facility_location <- tri_data %>% 
  select("facility_name", "street_address", "city", "county", "state", "zip_code", "bia_code", "tribal_land", "latitude", "longitude") %>% 
  distinct()

facility_location <- facility_location %>% 
  mutate(tribal_yes_no = case_when(is.na(tribal_land) == FALSE ~ "Tribal Land",
                                   TRUE ~ "Not Tribal Land"))

tri_data <- tri_data %>% 
  mutate(tribal_yes_no = case_when(is.na(tribal_land) == FALSE ~ "Tribal Land",
                                   TRUE ~ "Not Tribal Land"))

```

### Number of Facilities Owned by Each Company

The five most common parent companies are as follows: U.S. Department of Defense, CRH Americas Inc, Cemex Inc, Argos U.S. Corp, and Berkshire Hathaway Inc. The first listed is a government agency responsible for U.S. military protection, followed by four different building/construction/transport materials companies.

```{r number-of-facilities-by-company, echo = FALSE, message = FALSE, warning = FALSE}

tri_data %>% 
  select(standard_parent_co_name, parent_co_name, facility_name) %>% 
  group_by(standard_parent_co_name, parent_co_name) %>% 
  summarise(number_of_facilities = n_distinct(facility_name)) %>% 
  arrange(desc(number_of_facilities))

```

```{r number-of-facilities-by-state, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  group_by(state) %>% 
  count(state) %>% 
  arrange(desc(n))

tri_data %>% 
  select(facility_name, state) %>% 
  distinct(facility_name, state) %>% 
  group_by(state) %>%
  count(state) %>% 
  arrange(desc(n))
```

### Map of Facilities

The large majority of facilities in this dataset are located in North America, but a significant number of U.S.-owned facilities are strewn across the global map in other territories -- possibly shifting any inferences that can be made on the state of pollution solely in the U.S.

```{r facility-map, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "An interactive map displaying distinct U.S.-owned facility locations. Each point represents a facility and includes information such as facility name, parent company, industry sector, city, and state. 31,594 points are located in North America, and 127 are in other territories."}
tri_data %>% 
  distinct(longitude, latitude, facility_name, parent_co_name, industry_sector, city, state) %>% 
  leaflet() %>% 
  addTiles() %>% 
    addCircleMarkers(lng = ~longitude, 
                 lat = ~latitude, 
                 clusterOptions = markerClusterOptions(),
                 popup = ~paste0(facility_name, "<br>", parent_co_name, "<br>", industry_sector, "<br>", city, ", ", state))
```

### Percent of Facilities on Tribal Land

Though the percentage of facilities located on tribal land is less than 1% (there are `sum(facility_location$tribal_yes_no == "YES")` facilities on tribal land), and we are curious about any differences in the amount of chemicals used and released on tribal vs not tribal land.

```{r percent-tribal-land, echo = FALSE, message = FALSE, warning = FALSE}
facility_location %>% 
  select(tribal_yes_no) %>% 
  group_by(tribal_yes_no) %>%
  summarise(percent = 100 * n()/nrow(facility_location))

facility_location %>% 
  select(tribal_yes_no) %>% 
  group_by(tribal_yes_no) %>% 
  count(tribal_yes_no)
```

Using these tables, we have set up a groundwork map for placing chemical usage, which leads us to ask: knowing the drastic difference in percent of facilities on tribal land or not, does the location of tribal land have any effect on the type and amount of chemical release when compared to facilities not on tribal land?

We can take points from history and U.S. law, knowing that tribes of the American West often do have significant health hazards due to the many power plants, factories, mines, waste dumps, and testing sites that are placed on their land -- sometimes with permission as a means to acquiring much-needed funds, and sometimes out of a tribe's power. This question will appear further in visualizations.

## Industry Analysis

Within these observations of location, there is the question of each facility's purpose. Condensing the purpose of a factory into just 11 categories, their individual patterns of location, chemical use, and pollution can be tracked. This prevalence allows us to ask: is there a correlation between an industry's sector and their frequency of chemicals? Further, their frequency of specific types of chemicals -- carcinogens, PFAS', and PBT'?

```{r industry-sector-dataset, echo = FALSE, message = FALSE, warning = FALSE}
industry_info <- tri_data %>% 
  select(industry_sector, state, clean_air_act_chemical, carcinogen, metal_category, pbt, pfas, on_site_release_total, off_site_release_total, on_site_contained, off_site_contain) %>% 
  distinct(industry_sector, state, clean_air_act_chemical, carcinogen, metal_category, pbt, pfas, on_site_release_total, off_site_release_total, on_site_contained, off_site_contain)
```

```{r industry-sector-categories, echo = FALSE, message = FALSE, warning = FALSE}
tri_data <- tri_data %>% 
  mutate(industry_sector_category = case_when(industry_sector %in% c("Tobacco", "Beverages", "Food") ~ "Agriculture and Food Processing",
  industry_sector %in% c("Electrical Equipment", "Transportation Equipment", "Machinery", "Computers and Electronic     Products") ~ "Machinery and Technology",
  industry_sector %in% c("Miscellaneous Manufacturing", "Other") ~ "Misc",
  industry_sector %in% c("Chemicals", "Chemical Wholesalers", "Hazardous Waste") ~ "Chemicals and Waste",
  industry_sector %in% c("Leather", "Apparel", "Textiles", "Textile Product") ~ "Textiles and Leather",
  industry_sector %in% c("Petroleum Bulk Terminals", "Petroleum") ~ "Petroleum",
  industry_sector %in% c("Metal Mining", "Coal Mining") ~ "Mining",
  industry_sector %in% c("Nonmetallic Mineral Product", "Plastics and Rubber", "Fabricated Metals", "Primary Metals") ~ "Processed Materials",
  industry_sector %in% c("Furniture", "Wood Products", "Paper", "Printing", "Publishing") ~ "Wood and Paper",
  industry_sector %in% "Electric Utilities" ~ "Electric Utilities"))
```

### Industry Sector Distribution

The most common purpose of a U.S. facility is dealing with the production, waste, and wholesales of chemicals, accounting for almost 30% of U.S. industries as shown in the table. The least common accounting for about 0.4% is textile production which makes sense given trends in American imports.

```{r industry-sector-category-distribution, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A bar plot visualizes the distribution of industry sector categories in the Toxic Release Inventory (TRI). The plot displays the count of different industry sectors, with each bar representing a specific category. Chemical waste is the most prevalent sector, and textiles the least."}

tri_data %>% 
 select(industry_sector_category) %>% 
 group_by(industry_sector_category) %>%
 summarise(percent = 100 * n()/nrow(tri_data)) %>% 
 arrange(desc(percent))

tri_data %>% 
  ggplot(aes(y = fct_infreq(industry_sector_category), fill = industry_sector_category)) +
  geom_bar() +
  labs(title = "Industry Sector Prevalence",
    subtitle = "Distribution of Industry Sector Categories in the TRI",
    x = "Count",
    y = "Industry Sector Category",
    fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d()
```

### Carcinogen Percentage by Industry Sector Graph

Understandably fitting with the most common industry sector, the chemical-producing facilities (and processed materials) also use the most carcinogens (non-comprehensive), but the frequency seems to almost consistently grow with the prevalence of these sectors besides the food industry. That being said, every sector listed has a chemical usage that is majority non-carcinogenic.

```{r industry-sector-and-carcinogen-use, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A bar plot illustrates the distribution of carcinogen use across various industry sectors. Each bar represents an industry sector's prevalence, differentiated by the presence or absence of carcinogens, denoted by different fill colors. Every category uses mostly non-carcinogens, with the highest amount of carcinogens used in chemical and processing industries."}

tri_data %>% 
  ggplot(aes(y = fct_infreq(industry_sector_category), fill = carcinogen)) +
  geom_bar(position = "fill") +
  labs(title = "Carcinogen Use by Industry Sector",
       x = "Count",
       y = "Industry Sector",
       fill = "Carcinogenic") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)
```

## Chemical Analysis

Honing our observational questions down to just chemical trends of use and disposal now, we see how the laws of the EPA inform chemical procedures. Specifically in terms of the Clean Air Act, is there a differing correlation between the location and frequency of these specially-listed chemicals when compared to chemicals not listed on the Clean Air Act? And more broadly, is there any type, category, or closely-watched category of chemical that is most common despite the efforts of these laws?

```{r chemical-info-dataset, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info <- tri_data %>% 
  select(year, chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, chemical, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  arrange(chemical)
```

```{r chemical-count-by-year, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  select(chemical, year) %>% 
  distinct(chemical, year) %>% 
  group_by(year) %>%
  count(year) %>% 
  arrange(year)
```

The most commonly used chemical is lead, accounting for over 5% of use, followed by its compounds that make up almost 4% of use. Lead is known to cause poisoning of the human body, leading to neurological and reproductive health issues as well as impaired function of the kidneys and other body systems. Poisoning can be acquired through contaminated dust particles and products, contaminated drinking water, and contaminated soil.

```{r chemical-percent, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical) %>% 
  group_by(chemical) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))
```

### Chemical Category Calculations

\~\~ why are these separate from the corresponding graphs???? FIX

The large majority of chemicals used are not elemental metals or metals in general. About 63% are identified in the Clean Air Act, leaving 37% not present. Over 28% of chemicals used in these facilities are known to be carcinogenic, 17% are PBTs, and 0.02% are PFAS'. These smaller percentages are still a concern, and are in-part new variables to track. PFAS did not have to be reported to the TRI until 2020 (making this decade of data insufficient), even though over 2,000 facilities were known to be producing them before the regulation. Therefore, it has far fewer observations in comparison to other chemicals though used since the 1940s and known to cause certain cancers, impairment of the immune system, reproductive issues, and abnormal developmental effects. The reporting threshold for that particular subset of chemicals is 100 pounds per year.

```{r chemical-percent-elemental-metal-included, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(elemental_metal_included) %>% 
  group_by(elemental_metal_included) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))
```

```{r chemical-percent-pbt, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  select(pbt) %>% 
  group_by(pbt) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))
```

```{r chemical-percent-pfas, echo = FALSE, message = FALSE, warning = FALSE}
tri_data %>% 
  select(pfas) %>% 
  group_by(pfas) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))
```

#### Chemical Categories Graphs

```{r chemical-trends-by-category-elemental-metal-included, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot displays the prevalence of elemental metals used in U.S. facilities categorized by year from 2011 to 2021. Each bar represents the count of elemental metals with distinct colors indicating whether the chemicals used are or are not elemental metals. The large majority are not elemental metals, but the amount of elemental metals counted has been increasing slightly by year."}

tri_data %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>%
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = elemental_metal_included, fill = elemental_metal_included)) +
  geom_bar() +
  labs(title = "Elemental Metal Prevalence by Year",
       x = "Elemental Metal",
       y = "Count",
       fill = "Elemental Metal") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-percent-clean-air-act, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot displays the prevalence of Clean Air Act chemicals used in U.S. facilities categorized by year from 2011 to 2021. Each bar represents the count of chemicals with distinct colors indicating whether they are or are not identified in the Clean Air Act. The large majority are in the act, but the amount of non-act chemicals counted has been increasing slightly by year."}
 
chemical_info %>% 
  select(clean_air_act_chemical) %>% 
  group_by(clean_air_act_chemical) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = clean_air_act_chemical, fill = clean_air_act_chemical)) +
  geom_bar() +
  labs(title = "Clean Air Act Chemical Prevalence by Year",
       x = "Clean Air Act Chemical", 
       y = "Count",
       fill = "Clean Air Act Chemical") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year)
```

```{r chemical-percent-metal, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot displays the prevalence of metals used in U.S. facilities categorized by year from 2011 to 2021. Each bar represents the count of metals with distinct colors indicating whether the chemicals used are or are not metals. Most of them are not metals, and the amount of metals and non-metals counted has been increasing by year."}
 
chemical_info %>% 
  select(metal) %>% 
  group_by(metal) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))

tri_data %>%
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = metal, fill = metal)) +
  geom_bar() +
  labs(title = "Metal Prevalence by Year",
       x = "Metal",
       y = "Count",
       fill = "Metal") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-percent-metal-category, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A faceted bar plot depicts the prevalence of metal categories used by U.S. facilities from 2011 to 2021, whether a non-metal, metal with qualifiers, metal compound, chemical possibly containing metal, indivudually-listed metal compound, or elemental metal. Each bar represents the count of these metal categories, sorted by year, with varied colors denoting distinct metal category types."}

chemical_info %>% 
  select(metal_category) %>% 
  group_by(metal_category) %>%
  summarise(percent = 100 * n()/nrow(tri_data)) %>% 
  arrange(desc(percent))


# can i add line breaks into the y-axis values

tri_data %>%
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = metal_category, fill = metal_category)) +
  geom_bar() +
  labs(title = "Metal Category Prevalence by Year",
       x = "Metal Category",
       y = "Count",
       fill = "Metal Category") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~year) 
```

```{r chemical-trends-by-category-carcinogen, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot shows the prevalence of carcinogens used in U.S. facilities categorized by year from 2011 to 2021. Each bar represents the count of carcinogenic substances, sorted by year, with varied colors indicating different carcinogen categories. The majority of chemicals used are non-carcinogenic."}
chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = carcinogen, fill = carcinogen)) +
  geom_bar() +
  labs(title = "Carcinogen Prevalence by Year",
       x = "Carcinogenic",
       y = "Count",
       fill = "Carcinogenic") +
  scale_y_discrete(limits=c(0,10)) +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year)
```

```{r chemical-trends-by-category-pbt, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot shows the prevalence of PBTs used in U.S. facilities categorized by year from 2011 to 2021. Each bar represents the count of PBT substances, sorted by year, with different colors indicating whether a chemical is or is not a PBT. The large majority of chemicals used are non-PBTs, with the number of PBTs being relatively consistent by year."}

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = pbt, fill = pbt)) +
  geom_bar() +
  labs(title = "PBT Prevalence by Year",
       x = "PBT",
       y = "Count",
       fill = "PBT") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

```{r chemical-trends-by-category-pfas, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot shows the prevalence of PFAS chemicals used in U.S. facilities categorized by year from 2011 to 2021. Each bar represents the count of PFAS substances, sorted by year, with different colors indicating whether a chemical is or is not a PFAS. The overarching majority of chemicals used are not PFAS, and PFAS data only appears for the years 2020 and 2021."}

chemical_info %>% 
  select(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  distinct(year, elemental_metal_included, clean_air_act_chemical, metal, metal_category, carcinogen, pbt, pfas) %>% 
  ggplot(aes(x = pfas, fill = pfas)) +
  geom_bar() +
  labs(title = "PFAS Prevalence by Year",
       x = "PFAS",
       y = "Count",
       fill = "PFAS") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1) +
  facet_wrap(~year) 
```

##### Metal Category Pie Chart

```{r metal-category-pie-chart, echo=FALSE, message=FALSE, warning=FALSE, fig.alt = "A circular pie chart demonstrates the distribution of metal categories used by U.S. facilities in the dataset. Each segment of the pie represents a metal category with distinct colors, with the size proportional to the percentage of occurances. Non-metals make up 54% of the chart, and elemental metals make up 19% of the chart."}

tri_data %>% 
  group_by(metal_category) %>% 
  summarise(prop = n()/nrow(tri_data)) %>% 
  ggplot(aes(x = "", y = prop, fill = fct_inorder(metal_category))) +
  geom_col(color = "white", linewidth = .3) +
  geom_text(aes(x = "", y = prop, label = round(prop, 2)), position = position_stack(vjust = 0.5)) +
  labs(title = "Metal Category Distribution",
       x = "",
       y = "", 
       fill = "Metal Category") +
  theme_void() +
  coord_polar(theta = "y") +
  scale_fill_viridis_d(begin = 0.4, end = 0.9)
```

For above visualizations: elemental metals (some harmless, some causing neurotoxicity) keep a near-consistent distribution in the chemical data. The few increases this type of metal makes in the data is adjusted by the many additions of tracked chemicals in the dataframe, also causing the count for non-elemental-metals to increase. A similar state of increase goes for the amount of Clean Air Act chemicals and their growing amendments (still with more Clean Air Act chemicals than not) and with the general metals graph (more non-metals than metals, though both counts increasing). The amount of PBTs remains consistent with the non-PBTs increasing with new variables, and our two years of PFAS data show their consistency -- though, of course, not much judgement can be made through this small piece of information, and the same must go for much of these environmentally-altered graphs.

### Chemical Categories

There are `r sum(chemical_info$carcinogen == "YES")` carcinogens being observed -- a large number creating a small percentage.

```{r carcinogen-percent-yes, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, carcinogen) %>% 
  filter(carcinogen == "YES") %>% 
  group_by(chemical)
```

There are `r sum(chemical_info$carcinogen == "NO")` non-carcinogens being observed.

```{r carcinogen-percent-no, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, carcinogen) %>% 
  filter(carcinogen == "NO") %>% 
  group_by(chemical)
```

There are `r sum(chemical_info$pfas == "YES")` observations of PFAS chemicals in the dataset.

```{r pfas-percent-yes, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pfas) %>% 
  filter(pfas == "YES") %>% 
  group_by(chemical)
```

There are `r sum(chemical_info$pfas == "NO")` observations of non-PFAS chemicals in the dataset.

```{r pfas-percent-no, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pfas) %>% 
  filter(pfas == "NO") %>% 
  group_by(chemical)
```

There are `r sum(chemical_info$pbt == "YES")` observations of PBT chemicals in the dataset.

```{r pbt-percent-yes, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pbt) %>% 
  filter(pbt == "YES") %>% 
  group_by(chemical)
```

There are `r sum(chemical_info$pbt == "NO")` observations of non-PBT chemicals in the dataset.

```{r pbt-percent-no, echo = FALSE, message = FALSE, warning = FALSE}
chemical_info %>% 
  select(chemical, pbt) %>% 
  filter(pbt == "NO") %>% 
  group_by(chemical)
```

## Pollution analysis

From these distributions of observed chemicals, there is still the question of their disposal after use. With air, ground, and water being some possible points of disposal, we must assess the trends in these differing types of waste removal. In the past decade, what are these trends in pollution due to chemical release, whether on-site or off?

```{r pollution-info-dataset, echo = FALSE, message = FALSE, warning = FALSE}
pollution_info <- tri_data %>% 
  select(year, fugitive_air, stack_air, water, underground, underground_class_1, underground_class_2_through_5, landfills, land_treatment, surface_impoundment, other_disposal, on_site_release_total, public_treatment_total_transfer, off_site_release_total, off_site_recycled_total, off_site_energy_recovery_total, off_site_treated_total, total_transfer, total_releases, releases, on_site_contained, on_site_other, off_site_contain, off_site_other)
```

### On-Site Pollution by Category

#### By Type

On-site release types: variables `Underground`, `Landfill`, and `Surface Impoundment` are all zero.

```{r pollution-by-on-site-release-fugitive-air, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A smoothed scatter plot visualizes the relationship between fugitive air releases and on-site release totals of chemicals used in U.S. facilities from 2011 to 2021. Each point represents a data observation, and most years show a steep decline in fugitive air release with larger on-site release."}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = fugitive_air, color = fugitive_air, fill = fugitive_air)) + 
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Fugitive Air Release by Year",
       x = "On Site Release Total",
       y = "Fugitive Air") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-stack-air, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A smoothed scatter plot visualizes the relationship between stack air releases and on-site release totals of chemicals used in U.S. facilities from 2011 to 2021. Each point represents a data observation, and most years show a steep decline in stack air release with larger on-site release."}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = stack_air)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Stack Air Release by Year",
       x = "On Site Release Total",
       y = "Stack Air") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-water, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A smoothed scatter plot visualizes the relationship between water releases and on-site release totals of chemicals used in U.S. facilities from 2011 to 2021. Each point represents a data observation, and most years show a steep decline in water release with larger on-site release."}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = water)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Water Release by Year",
       x = "On Site Release Total",
       y = "Water") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-underground-class-1, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A smoothed scatter plot visualizes the relationship between underground (Class 1) releases and on-site release totals of chemicals used in U.S. facilities from 2011 to 2021. Each point represents a data observation, and most years show a steep decline in underground (Class 1) release with larger on-site release."}
 pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_1)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Underground Class 1 Releases by Year",
       x = "On Site Release Total",
       y = "Underground Class 1") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-underground-class-2-to-5, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A smoothed scatter plot visualizes the relationship between underground (Class 2-5) releases and on-site release totals of chemicals used in U.S. facilities from 2011 to 2021. Each point represents a data observation, and most years show a consistently straight line of underground (Class 2-5) release even with larger on-site release."}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_2_through_5)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Underground Class 2 Through 5 Release by Year",
       x = "On Site Release Total",
       y = "Underground Class 2 Through 5") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-land-treatment, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A smoothed scatter plot visualizes the relationship between land treatment and on-site release totals of chemicals used in U.S. facilities from 2011 to 2021. Each point represents a data observation, and most years show a small decline in land treatment with larger on-site release."}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = land_treatment)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Land Treatment by Year",
       x = "On Site Release Total",
       y = "Land Treatment") +
  theme_minimal() +
  facet_wrap(~ year)
```

```{r pollution-by-on-site-release-other-disposal, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A smoothed scatter plot visualizes the relationship between other disposal and on-site release totals of chemicals used in U.S. facilities from 2011 to 2021. Each point represents a data observation, and all years show a steep increase in other release with larger on-site release."}
pollution_info %>% 
  ggplot(aes(x = on_site_release_total, y = other_disposal)) +
  geom_smooth(color = "#00608E", linewidth = 0.6) +
  labs(title = "Other Disposal by Year",
       x = "On Site Release Total",
       y = "Other Disposal") +
  theme_minimal() +
  facet_wrap(~ year)
```

### Pollution by General Category (Including On-site-total)

Pollution variable `Releases` is all zeros.

```{r pollution-by-general-category, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in on-site release totals of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in on-site releases over the specified period. The on-site releases slowly decline over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = on_site_release_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "On Site Release Total by Year",
       x = "Year",
       y = "On Site Release Total") +
  theme_minimal()
```

```{r pollution-public-treatment-total-transfer, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in public treatment transfer totals of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these transfers over the specified period. The public treatment transfers slowly increase over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = public_treatment_total_transfer)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Transfer to Public Treatment by Year",
       x = "Year",
       y = "Total Transfer to Public Treatment") +
  theme_minimal()
```

```{r pollution-off-site-release-total, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in off-site release totals of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these releases over the specified period. The off-site release totals decrease over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_release_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Off Site Release Total by Year",
       x = "Year",
       y = "Off Site Release Total") +
  theme_minimal()
```

```{r pollution-off-site-recycled-total, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in off-site recycled totals of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these releases over the specified period. The off-site recycled totals only slightly increase by the end of the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_recycled_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Recycled Off Site by Year",
       x = "Year",
       y = "Total Recycled Off Site") +
  theme_minimal()
```

```{r pollution-off-site-energy-recovery-total, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in off-site energy recovery totals of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these energy recoveries over the specified period. The recovery totals drastically increase by the end of the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_energy_recovery_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Energy Recovery Off Site by Year",
       x = "Year",
       y = "Total Energy Recovery Off Site") +
  theme_minimal()
```

```{r pollution-off-site-treated-total, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in off-site treated totals of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these treated totals over the specified period. The off-site treated totals decrease by the end of the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_treated_total)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Treated Off Site by Year",
       x = "Year",
       y = "Total Treated Off Site") +
  theme_minimal()
```

```{r pollution-total-transfer, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in transfer totals of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these transfers over the specified period. The transfer totals increase by the end of the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = total_transfer)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Transfered by Year",
       x = "Year",
       y = "Total Transfered") +
  theme_minimal()
```

```{r pollution-total-releases, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in total releases of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these releases over the specified period. The total releases decrease over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = total_releases)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Total Released by Year",
       x = "Year", 
       y = "Total Released")
```

```{r pollution-on-site-contained, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in chemicals contained on-site from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these containments over the specified period. The total of on-site contained chemicals somewhat decrease over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = on_site_contained)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Amount Contained On Site By Year",
       x = "Year",
       y = "Amount Contained On Site") +
  theme_minimal()
```

```{r pollution-on-site-other, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in total on-site other disposal of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these disposals over the specified period. The total on-site disposals decrease consistently over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = on_site_other)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "On Site 'Other' by Year",
       x = "Year",
       y = "On Site 'Other'") +
  theme_minimal()
```

```{r pollution-off-site-contain, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in chemicals contained off-site from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these containments over the specified period. The total of off-site contained chemicals decrease over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_contain)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Contained Off Site by Year",
       x = "Year",
       y = "Contained Off Site") +
  theme_minimal()
```

```{r pollution-off-site-other, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A line plot visualizes the trend in total off-site other disposal of chemicals from U.S. facilities from 2011 to 2021. The line represents a smoothed curve depicting the change in these disposals over the specified period. The total off-site other disposals remain virtually consistent over the decade."}
pollution_info %>% 
  ggplot(aes(x = year, y = off_site_other)) + 
  geom_smooth(fill = "#00A790", color = "#00608E", linewidth = 0.6) +
  labs(title = "Off Site 'Other' by Year",
       x = "Year",
       y = "Off Site 'Other'") +
  theme_minimal()
```

```{r industry-sector-category-water-release, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot illustrates the water release across various industry sector categories for facilities in the U.S from 2011 to 2021. Each bar represents the quantity of water released, with distinct colors indicating different industry sector categories. Agriculture and food processing has the most water release of all industry sectors, and machinery has none."}
tri_data %>% 
  ggplot(aes(y = industry_sector_category, x = water, fill = industry_sector_category)) +
  geom_col() +
  labs(title = "Water Release by Industry Sector Category",
       x = "Water",
       y = "Industry Sector Category",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d()
```

# Detailed Analysis

```{r create-analysis-data-frame, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data <- tri_data %>% 
  filter(chemical %in% c("Arsenic", "Styrene", "Formaldehyde", "Nickel", "Chromium", "Lead", "Mercury", "Ammonia", "Phosphorus (yellow or white)", "Chloroform", "Benzo[g,h,i]perylene")) 

chemical_analysis_data <- chemical_analysis_data %>%
  mutate(off_site_total = off_site_release_total + off_site_recycled_total + off_site_energy_recovery_total + public_treatment_total_transfer + off_site_treated_total + total_transfer + off_site_contain + off_site_other)

chemical_analysis_data <- chemical_analysis_data %>% 
  mutate(on_site_total = fugitive_air + stack_air + water + underground + underground_class_1 + underground_class_2_through_5 + landfills + land_treatment + surface_impoundment + other_disposal + on_site_release_total + on_site_contained + on_site_other)
```

## Chemicals and Location

```{r count-chemical-and-type, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  count(chemical) %>% 
  arrange(desc(n))

chemical_analysis_data %>% 
  select(chemical, pfas, pbt, clean_air_act_chemical, carcinogen) %>% 
  group_by(chemical) %>% 
  count(pfas, pbt, clean_air_act_chemical, carcinogen) %>% 
  arrange(desc(n))
```

```{r chemical-use-by-facility, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  select(standard_parent_co_name, parent_co_name, facility_name, chemical) %>% 
  group_by(standard_parent_co_name, parent_co_name, facility_name) %>% 
  summarise(number_of_chemicals = n_distinct(chemical)) %>% 
  arrange(desc(number_of_chemicals))
```

```{r chemical-use-on-tribal-lands, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = "A grouped bar plot followed by a stacked bar plot visualizes the proportion of chemical use categorized by its presence on and off tribal lands from 2011 to 2021. Each bar segment represents the proportion of a chemical whether on or off tribal land, with color representing the specific chemical. Nickel, mercury, formaldehyde, chromium, chloroform, benzo[g,h,i]perylene, and arsenic are more apparent on tribal land than off. Phosphorus is not used at all on tribal land."}
chemical_analysis_data %>% 
  group_by(tribal_yes_no, chemical) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = chemical, y = prop, fill = tribal_yes_no)) +
  geom_col(position = "fill") +
  labs(title = "Chemical Use On and Off Tribal Lands",
       subtitle = "By Chemical",
       x = "Chemical",
       y = "Proportion",
       fill = "Chemical") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  select(chemical, tribal_yes_no) %>% 
  count(chemical, tribal_yes_no)

chemical_analysis_data %>% 
  ggplot(aes(y = tribal_yes_no, fill = chemical)) +
  geom_bar(position = "fill", color = "white") +
  labs(title = "Chemical Use On and Off Tribal Lands",
       subtitle = "By Tribal and Not Tribal Land",
       y = "Tribal Land?",
       x = "",
       fill = "Tribal Land?") +
  theme_minimal() +
  scale_fill_viridis_d()
```

```{r chemical-use-by-state, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>%
  select(state, chemical) %>%
  group_by(state) %>%
  summarise(chemical_use = n_distinct(chemical)) %>%
  arrange(desc(chemical_use))
```

### Overview graphs

## Chemical Release Graphs

```{r chemical-category-graphs, echo = FALSE, message = FALSE, warning = FALSE, fig.alt = c("For plot 1, a bar plot illustrates the frequency of eleven different chemicals used in U.S. faciltiies from 2011 to 2021 categorized by whether or not they are persistent, bioaccumulative, and toxic (PBT). Each bar represents a chemical, and the color distinction indicates whether the chemical falls under the PBT classification. Three of the eleven chosen chemicals are PBTs, and lead is the most commonly used PBT.", "For plot 2, a bar plot illustrates the frequency of eleven different chemicals used in U.S. faciltiies from 2011 to 2021 by whether or not they are categorized as PFAS. Each bar represents a chemical, and the color distinction indicates whether the chemical falls under the PFAS classification. None of the chosen chemicals are PFAS.", "For plot 3, a bar plot illustrates the frequency of eleven different chemicals used in U.S. faciltiies from 2011 to 2021 by whether or not they are categorized as carcinogenic. Each bar represents a chemical, and the color distinction indicates whether the chemical falls under the carcinogen classification. Six of the eleven chosen chemicals are carcinogens, with lead being the most commonly used carcinogen.", "For plot 4, a bar plot illustrates the frequency of eleven different chemicals used in U.S. faciltiies from 2011 to 2021 by whether or not they are classified as Clean Air Act chemicals. Each bar represents a chemical, and the color distinction indicates whether the chemical falls under the Act classification. Ammonia and lead are the only non-act chemicals in the chosen selection.")}
chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = pbt)) +
  geom_bar() +
  labs(title = "Is This Chemical a PBT?",
       y = "Chemical",
       fill = "PBT") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = pfas)) +
  geom_bar() +
  labs(title = "Is This Chemical a PFAS?",
       y = "Chemical",
       fill = "PFAS") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = carcinogen)) +
  geom_bar() +
  labs(title = "Is This Chemical Carcinogenic?",
       y = "Chemical",
       fill = "Carcinogenic") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)

chemical_analysis_data %>% 
  ggplot(aes(y = fct_infreq(chemical), fill = clean_air_act_chemical)) +
  geom_bar() +
  labs(title = "Is This Chemical Included in the Clean Air Act?",
       y = "Chemical",
       fill = "Clean Air Act Chemical?") +
  theme_minimal() +
  scale_fill_viridis_d(begin = 0.4, end = 1)
```

### Graphs of on-site release by type

```{r plot-fugitive-air-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = fugitive_air)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Fugitive Air in On Site Release Total",
       x = "On Site Release Total",
       y = "Fugitive Air",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-stack-air-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = stack_air)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Stack Air in On Site Release Total",
       x = "On Site Release Total",
       y = "Stack Air",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-water-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = water)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Water in On Site Release Total",
       x = "On Site Release Total",
       y = "Water",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-underground-class-1-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_1)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Underground Class 1 in On Site Release Total",
       x = "On Site Release Total",
       y = "Underground Class 1",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-underground-2-to-5-release-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = underground_class_2_through_5)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Underground Class 2 through 5 in On Site Release Total",
       x = "On Site Release Total",
       y = "Underground Class 2 through 5",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-land-treatment-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = land_treatment)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Land Treatment in On Site Release Total",
       x = "On Site Release Total",
       y = "Land Treatment",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-other-disposal-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = other_disposal)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "'Other' Disposal in On Site Release Total",
       x = "On Site Release Total",
       y = "'Other' Disposal",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-on-site-release-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_release_total, y = chemical, fill = chemical)) +
  geom_col() +
  labs(title = "On Site Release Total by Chemical",
       x = "On Site Release Total",
       y = "Chemical",
       fill = "Chemical") +
  theme_minimal() +
  scale_fill_viridis_d()
```

### Graphs of on-site disposal by type

```{r plot-on-site-contained-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_total, y = on_site_contained)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Contained On Site by Chemical",
       x = "On Site Total",
       y = "On Site Contained",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-on-site-other-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = on_site_total, y = on_site_other)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "On Site 'Other' and On Site Total",
       x = "On Site Total",
       y = "On Site 'Other'",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

### Graphs of off-site disposal by type

```{r plot-public-treatment-transfer-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = public_treatment_total_transfer)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Transfered to Public Treatment and Off Site Total",
       x = "Off Site Total",
       y = "Total Transfered to Public Treatment",
       color = "Chemical",
       alpha = "Saturation")  +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-release-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_release_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Off Site Release Total and Off Site Total",
       x = "Off Site Total",
       y = "Off Site Release Total",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-recycled-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_recycled_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Recycled off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Recycled Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-energy-recovery-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_energy_recovery_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Energy Recovered Off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Energy Recovered Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-treated-total-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_treated_total)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Treated Off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Treated Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-total-transfer-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = total_transfer)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Transfered and Off Site Total",
       x = "Off Site Total",
       y = "Total Transfered",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-total-releases-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = total_releases)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Released and Off Site Total",
       x = "Off Site Total",
       y = "Total Released",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-contain-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_contain)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Total Contained Off Site and Off Site Total",
       x = "Off Site Total",
       y = "Total Contained Off Site",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

```{r plot-off-site-other-by-chemical, echo = FALSE, message = FALSE, warning = FALSE}
chemical_analysis_data %>% 
  ggplot(aes(x = off_site_total, y = off_site_other)) +
  geom_point(aes(alpha = 0.3, color = chemical)) +
  labs(title = "Off Site 'Other' and Off Site Total",
       x = "Off Site Total",
       y = "Off Site 'Other'",
       color = "Chemical",
       alpha = "Saturation") +
  theme_minimal() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  facet_wrap(~ chemical)
```

### Chemical Release Trends by Year

Note: Phosphorus is not faceted by industry sector because there is only one industry that uses it.

```{r on-site-total-by-year}
chemical_analysis_data %>% 
  ggplot(aes(x = chemical, y = on_site_total, fill = chemical)) +
  geom_col(position = "dodge") +
  labs(title = "Total On Site By Chemical",
       subtitle = "On Site Release, Storage, and Disposal",
       x = "Chemical",
       y = "On Site Total",
       fill = "Chemical") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~ year)
```

```{r off-site-total-by-year}
chemical_analysis_data %>% 
  ggplot(aes(x = chemical, y = off_site_total, fill = chemical)) +
  geom_col(position = "dodge") +
    labs(title = "Total Off Site By Chemical",
       subtitle = "Off Site Release, Storage, Treatment, and Disposal",
       x = "Chemical",
       y = "Off Site Total",
       fill = "Chemical") +
  theme_minimal() +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~ year)
```

```{r arsenic-release-by-year}
chemical_analysis_data %>%
  filter(chemical == "Arsenic") %>%
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Arsenic Off Site Total",
       subtitle = "Arsenic Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r styrene-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Styrene") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Styrene Off Site Total",
       subtitle = "Styrene Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r formaldehyde-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Formaldehyde") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Formaldehyde Off Site Total",
       subtitle = "Formaldehyde Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r nickel-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Nickel") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Nickel Off Site Total",
       subtitle = "Nickel Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r chromium-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Chromium") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Chromium Off Site Total",
       subtitle = "Chromium Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)

chemical_analysis_data %>% 
  filter(chemical == "Chromium") %>% 
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Chromium On Site Total",
       subtitle = "Chromium On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r lead-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Lead") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Lead Off Site Total",
       subtitle = "Lead Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r mercury-release-by-year}
chemical_analysis_data %>%
  filter(chemical == "Mercury") %>%
  ggplot(aes(x = year, y = on_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Mercury On Site Total",
       subtitle = "Mercury On Site Release, Storage, and Disposal",
       x = "Year",
       y = "On Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r ammonia-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Ammonia") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Ammonia Off Site Total",
       subtitle = "Ammonia Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r phosphorus-release-by-year}
chemical_analysis_data %>%
  filter(chemical == "Phosphorus") %>%
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Phosphorus Off Site Total",
       subtitle = "Phosphorus Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
  fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d()
```

```{r chloroform-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Chloroform") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Chloroform Off Site Total",
       subtitle = "Chloroform Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

```{r benzo[g,h,i]perylene-release-by-year}
chemical_analysis_data %>% 
  filter(chemical == "Benzo[g,h,i]perylene") %>% 
  ggplot(aes(x = year, y = off_site_total, fill = industry_sector_category)) +
  geom_col(position = "dodge", just = 0.5) +
  labs(title = "Benzo[g,h,i]perylene Off Site Total",
       subtitle = "Benzo[g,h,i]perylene Off Site Release, Storage, Treatment, and Disposal",
       x = "Year",
       y = "Off Site Total",
       fill = "Industry Sector Category") +
  theme_minimal() +
  scale_fill_viridis_d() +
  facet_wrap(~ industry_sector_category)
```

# Conclusion

In recognizing the increase of observations, increase in pollution trends, and the environment of increasingly strict reporting laws, we must be careful not to make the assumption that pollution is truly on the rise, and that conditions are getting worse in general. What the legal environment shows is an increase in control and recognition of needed chemical regulations. What the work shows is a very actively polluted environment irregardless of these laws, and the continued prevalence of chemicals that are known to have been used as weapons of war and for body-preservation and alteration -- clearly harmful to human health. This is not to say that there is no improvement, but that there surely is much more that must be done.
